Dependency in Refinement Logics

### 2015-10-25 doc/paper.tex
Correct ctx wellformedness; add generalization to telescopes

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath, amsthm, stmaryrd, scalerel}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{proof}
\usepackage{supertabular}
\usepackage{unicode-math}
\usepackage{xltxtra}
\usepackage{xunicode}

@@

\newcommand\SortJdg{\mathsf{jdg}}
\newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
\newcommand\JdgValence[1]{\mathfrak{v}(#1)}
\newcommand\Metavars[1]{\mathfrak{M}(#1)}

@@

In order to properly treat generic judgments whose syntheses bind variables, we
will commit up front to using the abstract binding trees logical framework as a
universal domain of discourse, with a distinguished sort $\SortJdg$ for
object-judgments; and to each judgment $\JJ$ such that
$\IsWf{\Omega}{\Upsilon}{\Gamma}{\JJ}{\SortJdg}$, a \emph{valence}
$\JdgValence{\JJ}$ shall be associated which describes the binding structure of
its synthesis. Then, to each subgoal context $\Psi$, we shall associate a
metavariable context $\Metavars{\Psi}$ as follows:

\begin{align*}
  \IMode{\Metavars{\cdot}} &= \OMode{\cdot}\\
  \IMode{\Metavars{\Psi, \alpha:\JJ}} &= \OMode{\Metavars{\Psi}, \alpha : \JdgValence{\JJ}}
\end{align*}

Then, the judgments $\NoRefine{\JJ}{\tau}$ and $\Refine{\JJ}{\tau}{\Psi}{E}$
shall presuppose $\IsWf{\cdot}{\cdot}{\cdot}{\JJ}{\SortJdg}$, and the latter
shall postsuppose $\IsWf{\Metavars{\Psi}}{\cdot}{\cdot}{E}{\JdgValence{\JJ}}$.
When $E$ binds no variables or symbols, we shall write $E$ instead of
$\MkBTm{\cdot}{\cdot}{E}$.

%% In practice, this means that the
%% judgment $\Refine{\JJ}{\tau}{\Psi}{E}$ shall presuppose
%% $\IsWf{\cdot}{\cdot}{\cdot}{\JJ}{\SortJdg}$

@@

\begin{thm}
  The above rules all wellformed definitions of tactics.
  \begin{enumerate}
    \item $\IsTac{\IdTac}$.
    \item $\IsTac{\FailTac}$.
    \item If $\IsTac{\tau_1}$ and $\IsTac{\tau_2}$, then
      $\IsTac{\OrElseTac{\tau_1}{\tau_2}}$.

    \item If $\IsTac{\tau_1}$ and for any $\Member{\alpha}{I}$,
      $\IsTac{\tau_\alpha}$, then
      $\IsTac{\ThenTac{\tau_1}{\MkFam{\tau}{\alpha}{I}}}$.
  \end{enumerate}
\end{thm}
\begin{proof}
  Disjointness for each tactical is immediately evident; it suffices, then, to
  verify the postsupposition
  $\IsWf{\Metavars{\Psi}}{\cdot}{\cdot}{E}{\JdgValence{\JJ}}$ in each case.
  \begin{enumerate}
    \item[(1--2)] Immediate.
    \item[(3)]
      By induction on derivations of
      $\Refine{\JJ}{\OrElseTac{\tau_1}{\tau_1}}{E}{\Psi}$.
      \begin{description}
        \item[Case $\Refine{\JJ}{\tau_1}{\Psi}{E}$.] Validity follows from the
          inductive hypothesis $\IsTac{\tau_1}$.
        \item[Case $\NoRefine{\JJ}{\tau_1}$.] Validity follows from the
          inductive hypothesis $\IsTac{\tau_2}$.
      \end{description}

@@

We can now encode \emph{PairIntro} with a single refinement rule,
$\DPairIntroRuleNullary$:

\[
  \Refine{\IsTrueUnmoded{\TyDPair{A}{x}{B}}}{\DPairIntroRuleNullary}{%
    \left\{\begin{array}{l}
        \OSG{\alpha}{\IsTrueUnmoded{A}}\\
        \OSG{\beta}{\IsTrueUnmoded{\Subst{\alpha}{x}{B}}}
    \end{array}\right.
  }{
    \Tuple{\alpha,\beta}
  }
\]

All that remains is to give a new definition of the sequencing tactical
$\ThenTac{\tau_1}{\MkFam{\tau}{\alpha}{I}}$ which accounts for this dependency.
To this end, we will need to implement an auxiliary judgment
\framebox{$\ThenMultiAux{\MkFam{\tau}{\alpha}{I}}{E}{\Phi}{E'}{\Phi'}$} that
allows us to apply the family of tactics $\MkFam{\tau}{\alpha}{I}$ pointwise to
a proof state, simultaneously propagating refinements rightward through
substitution. This judgment, which presupposes $\IsEquiv{I}{\Dom\Phi}$, is
defined by recursion on the telescope $\Phi$, viewed as a \emph{cons}-list:

>>>>

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath, amssymb, amsthm, stmaryrd, scalerel}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{proof}
\usepackage{supertabular}
\usepackage{unicode-math}
\usepackage{xltxtra}
\usepackage{xunicode}

@@

\newcommand\IsCtx[2]{
  \IMode{#1}\ \textit{ctx}\leadsto \OMode{#2}
}
\newcommand\IsTele[2]{
  \IMode{#1}\ \textit{tele}\leadsto \OMode{#2}
}

\newcommand\SortJdg{\mathsf{jdg}}
\newcommand\MkBTm[3]{\{#1\}[#2].\, #3}
\newcommand\JdgValence[1]{\mathfrak{v}(#1)}

@@

In order to properly treat generic judgments whose syntheses bind variables, we
will commit up front to using the abstract binding trees logical framework as a
universal domain of discourse, with a distinguished sort $\SortJdg$ for
object-judgments; and to each judgment $\JJ$ such that
$\IsWf{\Omega}{\Upsilon}{\Gamma}{\JJ}{\SortJdg}$, a \emph{valence}
$\JdgValence{\JJ}$ shall be associated which describes the binding structure of
its synthesis.

Then, we will need to define the wellformedness of subgoal contexts $\Psi$,
simultaneously associating to $\Psi$ a metavariable context $\Omega$ which will
stand for the collected syntheses of $\Psi$:
\begin{gather*}
  \infer{
    \IsCtx{\cdot}{\cdot}
  }{
  }\qquad
  \infer{
    \IsCtx{\Psi,\alpha:\JJ}{\Omega,\alpha:\JdgValence{\JJ}}
  }{
    \IsCtx{\Psi}{\Omega} &
    \IsWf{\cdot}{\cdot}{\cdot}{\JJ}{\SortJdg}
  }
\end{gather*}

\begin{remark}
  Note how the context-wellformedness judgment artificially requires that $\JJ$
  be wellformed in the empty metavariable context.
\end{remark}

Then, the judgments $\NoRefine{\JJ}{\tau}$ and $\Refine{\JJ}{\tau}{\Psi}{E}$
shall presuppose $\IsWf{\cdot}{\cdot}{\cdot}{\JJ}{\SortJdg}$, and the latter
shall postsuppose $\IsCtx{\Psi}{\Omega}$ and $\IsWf{\Omega}{\cdot}{\cdot}{E}{\JdgValence{\JJ}}$.
When $E$ binds no variables or symbols, we shall write $E$ instead of
$\MkBTm{\cdot}{\cdot}{E}$.

@@

\begin{thm}
  The above rules all define valid tactics.
  \begin{enumerate}
    \item $\IsTac{\IdTac}$.
    \item $\IsTac{\FailTac}$.
    \item If $\IsTac{\tau_1}$ and $\IsTac{\tau_2}$, then
      $\IsTac{\OrElseTac{\tau_1}{\tau_2}}$.

    \item If $\IsTac{\tau_1}$ and for any $\Member{\alpha}{I}$,
      $\IsTac{\tau_\alpha}$, then
      $\IsTac{\ThenTac{\tau_1}{\MkFam{\tau}{\alpha}{I}}}$.
  \end{enumerate}
\end{thm}
\begin{proof}
  Disjointness for each tactical is immediately evident; it suffices, then, to
  verify the postsuppositions $\IsCtx{\Psi}{\Omega}$ and
  $\IsWf{\Omega}{\cdot}{\cdot}{E}{\JdgValence{\JJ}}$ in each case.
  \begin{enumerate}
    \item[(1--2)] Immediate.
    \item[(3)]
      By induction on derivations of
      $\Refine{\JJ}{\OrElseTac{\tau_1}{\tau_1}}{E}{\Psi}$.
      \begin{description}
        \item[Case $\Refine{\JJ}{\tau_1}{\Psi}{E}$.] Validity follows from the
          inductive hypothesis $\IsTac{\tau_1}$.
        \item[Case $\NoRefine{\JJ}{\tau_1}$.] Validity follows from the
          inductive hypothesis $\IsTac{\tau_2}$.
      \end{description}

@@

Briefly, we can adjust the postsuppositions of the refinement judgment to
account for this generalization, by first characterizing the wellformed
telescopes:
\begin{gather*}
  \infer{
    \IsTele{\cdot}{\cdot}
  }{
  }\qquad
  \infer{
    \IsTele{\Psi,\alpha:\JJ}{\Omega,\alpha:\JdgValence{\JJ}}
  }{
    \IsTele{\Psi}{\Omega} &
    \IsWf{\Omega}{\cdot}{\cdot}{\JJ}{\SortJdg}
  }
\end{gather*}

Now, $\Refine{\JJ}{\tau}{\Psi}{E}$ shall postsuppose $\IsTele{\Psi}{\Omega}$
and $\IsWf{\Omega}{\cdot}{\cdot}{E}{\JdgValence{\JJ}}$.

\subsection{Dependent Tacticals}
All that remains is to give a new definition of the sequencing tactical
$\ThenTac{\tau_1}{\MkFam{\tau}{\alpha}{I}}$ which accounts for this dependency.
To this end, we will need to implement an auxiliary judgment
\framebox{$\ThenMultiAux{\MkFam{\tau}{\alpha}{I}}{E}{\Phi}{E'}{\Phi'}$} that
allows us to apply the family of tactics $\MkFam{\tau}{\alpha}{I}$ pointwise to
a proof state, simultaneously propagating refinements rightward through
substitution. This judgment, which presupposes $\IsEquiv{I}{\Dom\Phi}$, is
defined by recursion on the telescope $\Phi$, viewed as a \emph{cons}-list:
